---
title: "Point ragdoc at your docs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Point ragdoc at your docs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

`ragdoc` turns your documentation into a retrieval tool a coding agent can call.
Instead of guessing, the agent retrieves the *actual* passages from your docs,
each returned with a `[source · url]` citation it can quote back.

The whole workflow is three steps: describe the sources, build the store once,
then query it. The chunks below are not evaluated (building needs network access
and an API key), but they are the real calls.

## 1. Describe your sources

A source is a site to crawl. Give it a short `name` (used for citations and for
filtering later) and, optionally, a `pattern` regex so you ingest only the pages
you want and skip navigation and asset links.

```{r}
library(ragdoc)

src <- sources(
  web("handbook", "https://docs.example.com/", pattern = "\\.html$")
)
```

You can list several sources in one spec; their names must be unique.

```{r}
src <- sources(
  web("handbook", "https://docs.example.com/", pattern = "\\.html$"),
  web("api",      "https://api.example.com/",  pattern = "\\.html$")
)
```

### Local Markdown directories

A source can also be a directory of Markdown files on disk. `local_dir()` walks
the directory (recursively by default), ingesting files whose names match
`pattern` — `"\\.md$"` by default. Each file's path becomes its citation
locator, so retrieved passages point back at the file they came from.

```{r}
src <- sources(
  local_dir("notes", "~/project/docs")
)
```

Tune the walk with `pattern` and `recursive`, and mix local and web sources in
one spec:

```{r}
src <- sources(
  web("api",   "https://api.example.com/", pattern = "\\.html$"),
  local_dir("notes", "~/project/docs", pattern = "\\.(md|markdown)$", recursive = FALSE)
)
```

`local_dir()` is a provisional name (the natural peer of `web()`, `local()`,
collides with base R's `local()`), so it may change before a stable release.

## 2. Build the store

`build_store()` crawls each source, converts pages to markdown, chunks and
embeds them, and writes a DuckDB store with a hybrid (vector + keyword) index.
This is the one-time, network-bound step. By default it embeds with OpenAI's
`text-embedding-3-small`, so set `OPENAI_API_KEY` in `~/.Renviron` first — or
pass your own `embed=` function.

```{r}
build_store(src, "handbook.duckdb")
```

A broken page or an unreachable source is logged and skipped, not fatal, so you
always get a usable store.

## 3. Query it

Open the store and search. `search_docs()` returns a single formatted string —
the passages, each cited with its source and URL — which is exactly what you
hand to an agent.

```{r}
store <- connect_store("handbook.duckdb")
cat(search_docs(store, "how do I rotate the signing keys", n = 5))
```

Restrict to one source by name:

```{r}
cat(search_docs(store, "rate limits", source = "api"))
```

`connect_store()` returns a plain `ragnar` store, so you can drop down to
`ragnar` directly any time you need to.

## Serving it to an agent over MCP

The point of all this is to expose the store to a coding agent. `serve_mcp()`
wraps the store in a single Model Context Protocol tool and starts a server.
Put this in a script and point your agent at it:

```{r}
library(ragdoc)
serve_mcp("handbook.duckdb",
  name        = "search_handbook",
  description = "Search the Example Co engineering handbook.")
```

The agent now has a `search_handbook` tool. Pass `sources =` to hint at the
valid `source` filter values in the tool description. If the store has not been
built yet, the server still starts and the tool says so rather than crashing —
convenient while you are still setting things up.
