---
title: "Point ragdoc at your docs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Point ragdoc at your docs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

`ragdoc` turns your documentation into a retrieval tool a coding agent can call.
Instead of guessing, the agent retrieves the *actual* passages from your docs,
each returned with a `[source · url]` citation it can quote back.

The whole workflow is three steps: describe the sources, build the store once,
then query it. The chunks below are not evaluated (building needs network access
and an API key), but they are the real calls.

## 1. Describe your sources

A source is a site to crawl. Give it a short `name` (used for citations and for
filtering later) and, optionally, a `pattern` regex so you ingest only the pages
you want and skip navigation and asset links.

```{r}
library(ragdoc)

src <- sources(
  web("handbook", "https://docs.example.com/", pattern = "\\.html$")
)
```

You can list several sources in one spec; their names must be unique.

```{r}
src <- sources(
  web("handbook", "https://docs.example.com/", pattern = "\\.html$"),
  web("api",      "https://api.example.com/",  pattern = "\\.html$")
)
```

### Local directories of files

A source can also be a directory of files on disk. `local_dir()` walks the
directory (recursively by default) and ingests every file, reading each by its
type: plain text and Markdown verbatim, source code (`.R`, `.py`, `.qmd`, …)
wrapped in a language-fenced block, and everything else (`.pdf`, `.docx`,
`.pptx`, `.html`, …) converted with MarkItDown via `ragnar::read_as_markdown()`.
Each file's path becomes its citation locator, so retrieved passages point back
at the file they came from.

```{r}
src <- sources(
  local_dir("notes", "~/project/docs")
)
```

Narrow the walk with `pattern` (a regex on file names) and `recursive`, and mix
local and web sources in one spec:

```{r}
src <- sources(
  web("api",   "https://api.example.com/", pattern = "\\.html$"),
  local_dir("notes", "~/project/docs", pattern = "\\.(md|qmd|pdf)$", recursive = FALSE)
)
```

Converting non-text files needs a Python environment with `markitdown`
installed; plain text, Markdown, and code ingest without it. Local files over
25 MB are skipped (large PDFs can exhaust memory).

### Sources from a YAML file

To keep the corpus in a config file instead of R code, list it in a
`sources.yml` and load it — an entry with a `root_url` becomes a `web()` source,
one with a `path` a `local_dir()`:

```yaml
- name: api
  root_url: https://api.example.com/
  crawl_pattern: "\\.html$"
- name: notes
  path: ~/project/docs
```

```{r}
src <- load_sources("sources.yml")
```

## 2. Build the store

`build_store()` crawls each source, converts pages to markdown, chunks and
embeds them, and writes a DuckDB store with a hybrid (vector + keyword) index.
This is the one-time, network-bound step. By default it embeds with OpenAI's
`text-embedding-3-small`, so set `OPENAI_API_KEY` in `~/.Renviron` first — or
pass your own `embed=` function.

```{r}
build_store(src, "handbook.duckdb")
```

A broken page or an unreachable source is logged and skipped, not fatal, so you
always get a usable store.

When your sources change, you don't have to pay for a full rebuild. `refresh_store()`
re-ingests into the existing store and re-embeds only what is new or edited,
skipping everything unchanged:

```{r}
refresh_store(src, "handbook.duckdb")
```

Refresh is add-and-update only — it never removes anything, so reach for a full
rebuild (`build_store(overwrite = TRUE)`) when pages are removed or renamed, or
when you change the chunking or the embedder.

## 3. Query it

Open the store and search. `search_docs()` returns a single formatted string —
the passages, each cited with its source and URL — which is exactly what you
hand to an agent.

```{r}
store <- connect_store("handbook.duckdb")
cat(search_docs(store, "how do I rotate the signing keys", n = 5))
```

Restrict to one source by name:

```{r}
cat(search_docs(store, "rate limits", source = "api"))
```

`connect_store()` returns a plain `ragnar` store, so you can drop down to
`ragnar` directly any time you need to.

## Serving it to an agent over MCP

The point of all this is to expose the store to a coding agent. `serve_mcp()`
wraps the store in a single Model Context Protocol tool and starts a server.
Put this in a script and point your agent at it:

```{r}
library(ragdoc)
serve_mcp("handbook.duckdb",
  name        = "search_handbook",
  description = "Search the Example Co engineering handbook.")
```

The agent now has a `search_handbook` tool. Pass `sources =` to hint at the
valid `source` filter values in the tool description. If the store has not been
built yet, the server still starts and the tool says so rather than crashing —
convenient while you are still setting things up.
